---
title: "Norwegian unemployment data"
author: "Lars Raaum"
date: "1/23/2021"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(shiny)
library(dplyr)
library(ggplot2)
library(httr)
library(rjstat)

```

## Norwegian unemployment data
- This is a simple app visualizing Norwegian unemployment data from the Norwegian Labour Force Survey
- The Survey is done quarterly among a large sample of the Norwegian population
- The most recent data is from Q3 2020, thus i compare Q3 data for the entire time series
- The app lets you explore the dataset in dimensions of time and age groups

## Getting the data
- The data is collected from the Statistics Norway (SSB) API and some minor cleaning operations are performed 
```{r data}
options(encoding="UTF-8")
url <- "https://data.ssb.no/api/v0/no/table/05110/"

data <- '
{
  "query": [
    {
      "code": "ArbStyrkStatus",
      "selection": {
        "filter": "item",
        "values": [
          "2"
        ]
      }
    },
    {
      "code": "Alder",
      "selection": {
        "filter": "item",
        "values": [
          "15-74",
          "15-24",
          "25-54",
          "55-74"
        ]
      }
    },
    {
      "code": "ContentsCode",
      "selection": {
        "filter": "item",
        "values": [
          "Prosent"
        ]
      }
    },
    {
      "code": "Tid",
      "selection": {
        "filter": "item",
        "values": [
          "1988K3",
          "1989K3",
          "1990K3",
          "1991K3",
          "1992K3",
          "1993K3",
          "1994K3",
          "1995K3",
          "1996K3",
          "1997K3",
          "1998K3",
          "1999K3",
          "2000K3",
          "2001K3",
          "2002K3",
          "2003K3",
          "2004K3",
          "2005K3",
          "2006K3",
          "2006K3 Gml",
          "2007K3",
          "2008K3",
          "2009K3",
          "2010K3",
          "2011K3",
          "2012K3",
          "2013K3",
          "2014K3",
          "2015K3",
          "2016K3",
          "2017K3",
          "2018K3",
          "2019K3",
          "2020K3"
        ]
      }
    }
  ],
  "response": {
    "format": "json-stat2"
  }
}'
    d.tmp <- POST(url , body = data, encode = "json", verbose())
    # Extract contents from d.tmp as text, then fromJSONstat
    sbtabell <- fromJSONstat(content(d.tmp, "text"))
    
    sbtabell$alder <- as.factor(sbtabell$alder)
    sbtabell$value <- as.numeric(sbtabell$value)
    sbtabell$kvartal <- substr(sbtabell$kvartal, 0, 4)
    sbtabell$kvartal <- as.numeric(sbtabell$kvartal)
```

## Presenting the data in a Shiny app
- A slider is available for the range of years you want to show in the graph
- The graph shows the unemployment rate for Q3 in each year, by age group.
```{r input}

        ggplot(sbtabell, aes(x=kvartal, y=value)) + geom_line(aes(colour=alder, group = alder))
        

```

## Further links
The full application is available at https://lraaum.shinyapps.io/Week4/ 